Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-07-14 writeup/rocha_etal.tex
Fixes small typos

\section{The LLC numerical simulations}
We use the output of two latitude-longitude polar cap (LLC)
realistic numerical simulations. The outputs
analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
numerical solutions on a LLC grid  \citep{forget_etal2015} with
90-vertical-levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011. Both simulations were forced by
tides and 6-hourly surface atmospheric fields. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to September 2012.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016}.  Details of the LLC simulations are provided in the supplemental material.

@@

\section{Statistics of the surface lateral velocity tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme. We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
These diagnostics highlight the submesoscale structures in the flow
\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.

@@

\section{Summary and Conclusion}
Our work adds to recent studies that present modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
 \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
enhancement of inertia-gravity waves due to re-stratification of
the upper ocean.

\cite{dasaro1978} showed that the velocity of linear internal waves
in the mixed layer strongly depends on the density jump at the mixed-layer
base, with largest mixed layer velocities when the jump is strongest.
In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally. In particular the baroclinic modes are
significantly more surface intensified in late summer/early fall
 (supplemental material).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore one expects a strong seasonality in the near-surface
expression of internal tides and
other submesoscale inertia-gravity waves generated through internal tide interactions.
Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
those waves project on large horizontal scales \citep[e.g, ][]{qi_etal1995}.

>>>>

\section{The LLC numerical simulations}
We use the output of two latitude-longitude polar cap (LLC)
realistic numerical simulations. The outputs
analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
numerical solutions on a LLC grid  \citep{forget_etal2015} with
90-vertical-levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011. Both simulations were forced by
tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to September 2012.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016}.  Details of the LLC simulations are provided in the supplemental material.

@@

\section{Statistics of the surface lateral velocity gradient tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme. We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
These diagnostics highlight the submesoscale structures in the flow
\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.

@@

\section{Summary and Conclusion}
Our work adds to recent modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
 \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
enhancement of inertia-gravity waves due to re-stratification of
the upper ocean.

\cite{dasaro1978} showed that the velocity of linear internal waves
in the mixed layer strongly depends on the density jump at the mixed-layer
base, with largest mixed layer velocities when the jump is strongest.
In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally. In particular the baroclinic modes are
significantly more surface intensified in late summer/early fall
 (supplemental material).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore one expects a strong seasonality in the near-surface
expression of internal tides and
other submesoscale inertia-gravity waves generated through internal tide interactions.
Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
those waves project onto large horizontal scales \citep[e.g, ][]{qi_etal1995}.
